---
title: "Varroa re-mapped"
author: "Lazzat Aibekova"
date: "6/04/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
```

```{r library}
library(tidyverse)
library(sleuth)
```


```{r sleuth}
samples <- tibble(id = dir(file.path("mondet/kallisto_mite"))) %>% mutate(path = paste0("mondet/kallisto_mite/", id)) %>%
  left_join(read_tsv("mondet/treatments.tsv", col_types =  cols())) %>% mutate(sample = id, condition = shortName)
```


### Comparing males vs young

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("young", "males")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)


```
```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionyoung')
results_table = sleuth_results(so, 'conditionyoung')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```

###Young vs Phoretic

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("young", "phoretic")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionyoung')
results_table = sleuth_results(so, 'conditionyoung')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```

### Phoretics vs caged

There were nio

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("phoretic", "phoreticCaged")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionphoreticCaged')
results_table = sleuth_results(so, 'conditionphoreticCaged')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```
### Phoretic vs arresting

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("phoretic", "arresting")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```
```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionphoretic')
results_table = sleuth_results(so, 'conditionphoretic')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```

Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```
### Аrresting vs. pre-laying

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("arresting", "prelaying")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionprelaying')
results_table = sleuth_results(so, 'conditionprelaying')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```
### Pre-laying vs. Laying

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("prelaying", "laying")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionprelaying')
results_table = sleuth_results(so, 'conditionprelaying')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```
### Laying vs. post-laying

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("laying", "postlaying")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionpostlaying')
results_table = sleuth_results(so, 'conditionpostlaying')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```

### Post-laying vs. Emerging

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("postlaying", "emerging")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)

```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionpostlaying')
results_table = sleuth_results(so, 'conditionpostlaying')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```

### Post-laying vs. non-laying

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("postlaying", "nonlaying")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionpostlaying')
results_table = sleuth_results(so, 'conditionpostlaying')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```

### Laying vs. non-laying

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("laying", "nonlaying")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionnonlaying')
results_table = sleuth_results(so, 'conditionnonlaying')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```

### Emerging vs. Phoretic

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("emerging", "phoretic")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionphoretic')
results_table = sleuth_results(so, 'conditionphoretic')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```

### Emerging vs. Young

```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("emerging", "young")), extra_bootstrap_summary = TRUE)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
sleuth_table <- sleuth_results(so, 'reduced:full', 'lrt', show_all = FALSE)
sleuth_significant <- dplyr::filter(sleuth_table, qval <= 0.05)
dim(sleuth_significant)
head(sleuth_significant)
```

```{r}
models(so)
```
Number of Differentially expressed genes (TRUE):
```{r}
so = sleuth_wt(so, 'conditionyoung')
results_table = sleuth_results(so, 'conditionyoung')
Number_DE_genes = table(results_table$qval <= 0.05)
Number_DE_genes
```
Number of up-regulated (TRUE) and down-regulated (FALSE) genes:
```{r}
results_table
DE_genes = subset(results_table, qval <= 0.05)
DE_genes
up_down_regulated = table(DE_genes$b > 0)
up_down_regulated

```