---
title: "GOterm enrichment analysis"
author: "Lazzat Aibekova"
date: "June 11, 2019"
output:
  html_document: default
    toc: true
    toc_depth: 3
    number_section: 
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(topGO)
library(sleuth)
library(tidyverse)

```

```{r sleuth}
samples <- tibble(id = dir(file.path("mondet/kallisto_mite"))) %>% mutate(path = paste0("mondet/kallisto_mite/", id)) %>%
  left_join(read_tsv("mondet/treatments.tsv", col_types =  cols())) %>% mutate(sample = id, condition = shortName)
```

#transcript to gene table
```{r}
t2g = read.csv("genebank_gene_map.csv", header=T)
```

###LOOP STARTS HERE
```{r}
so <- sleuth_prep(dplyr::filter(samples, shortName %in% c("young", "males")), target_mapping = t2g,aggregation_column = 'ens_gene', gene_mode = T)
so <- sleuth_fit(so, ~condition, 'full')
so <- sleuth_fit(so, ~1, 'reduced')
so <- sleuth_lrt(so, 'reduced', 'full')
so = sleuth_wt(so, 'conditionyoung')
results_table = sleuth_results(so, 'conditionyoung')
DE_genes = subset(results_table, qval <= 0.05)
```
#Table with up-regulated genes
```{r}
up_regulated = subset(DE_genes, b > 0) #beta postive is up-regulated
```

#Table with down-regulated genes
```{r}
down_regulated = subset(DE_genes, b < 0) #beta postive is up-regulated
```

###Preapre inputs for sampleOGData of class topGOdata
##Males vs Young
```{r}
geneList = up_regulated[,2]
names(geneList) = as.character(up_regulated[,1])
geneList = sort(geneList, decreasing = TRUE)
```
#This Function is to build gene2GO map for topGO
```{r}
readMappings2 = function (file, sep = ",", IDsep = "\t") 
{
  b <- read.delim(file = file, header = TRUE, row.names = 1, quote = "", 
                  sep = sep, colClasses = "character")
  lst <- by(b[1:length(b)], rownames(b), as.character, simplify = TRUE)
  lst2 = lapply(lst, function(x) x[x != ""])
  return(lst2)
}
```

```{r}
gene2GO = readMappings2("GOterm/vdesgoassoc.csv") #gene if to GO id map
```
#Function selects genes with p value < 0.05
```{r}
topDiffGenes = function (allScore) {
    return(allScore < 0.05)
}
```

```{r}
 sampleGOdata <- new("topGOdata", description = "Simple session", ontology = "BP", allGenes = geneList, geneSel = topDiffGenes,  nodeSize = 5, annot = annFUN.gene2GO, gene2GO = gene2GO)
```

```{r}
sampleGOdata
```
###Fisher's exact test - based on gene counts
```{r}
resultFisher <- runTest(sampleGOdata, algorithm = "classic", statistic = "fisher")

```

```{r}
resultFisher
```
###Kolmogorov_smirnov - based on gene scores
```{r}
resultKS <- runTest(sampleGOdata, algorithm = "classic", statistic = "ks")
```

```{r}
resultKS.elim <- runTest(sampleGOdata, algorithm = "elim", statistic = "ks")
```
###Analyse Results
```{r}
allRes <- GenTable(sampleGOdata, classicFisher = resultFisher, classicKS = resultKS, elimKS = resultKS.elim, orderBy = "elimKS", ranksOf = "classicFisher", topNodes = 10)
```
#Function to map colors
```{r}
colMap <- function(x) {
  .col <- rep(rev(heat.colors(length(unique(x)))), time = table(x))
  return(.col[match(1:length(x), order(x))])
}
```

```{r}
pValue.classic <- score(resultKS)
pValue.elim <- score(resultKS.elim)[names(pValue.classic)]
gstat <- termStat(sampleGOdata, names(pValue.classic))
gSize <- gstat$Annotated / max(gstat$Annotated) * 4
gCol <- colMap(gstat$Significant)
plot(pValue.classic, pValue.elim, xlab = "p-value classic", ylab = "p-value elim",
pch = 19, cex = gSize, col = gCol)
```

```{r}
showSigOfNodes(sampleGOdata, score(resultKS.elim), firstSigNodes = 5, useInfo = 'all')
```

